---
title: "Sick dataset analysis"
author: "Bogdan Jastrzębski"
date: "26 kwietnia 2020 r."
output:
  bookdown::pdf_book:
    number_sections: TRUE
    toc: true
    fig_caption: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, echo = FALSE, message=FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(OpenML)
library(dplyr)
library(tidyverse)
library(kableExtra)
library(visdat)
library(naniar)
library(DataExplorer)
library(DALEX)
library(GGally)
library(gridExtra)
library(auprc)
library(e1071)
library(rpart.plot)
library(imputeTS)
```


```{r data, echo = FALSE, message=FALSE, warning = FALSE, include = FALSE}
set.seed(10)
sick <- getOMLDataSet(data.name = "sick")$data

I <- read.table("indeksy_treningowe.txt")
i <- I$x

sick_tidy <- sick %>%
  dplyr::select(-TBG,
         -TBG_measured,
         -FTI_measured,
         -T4U_measured,
         -TT4_measured,
         -T3_measured,
         -TSH_measured)

sick_tidy <- sick_tidy %>% 
  dplyr::select(-hypopituitary)

ind <- 1:nrow(sick_tidy) %in% i

sick_tidy <- cbind(sick_tidy, ind)

sick_tidy <- sick_tidy %>% na.exclude()

sick_t <- sick_tidy %>%
  mutate(log_TSH = log(TSH),
         sqrt_T3 = sqrt(T3),
         sqrt_TT4 = sqrt(TT4),
         log_T4U = log(T4U),
         sqrt_FTI = sqrt(FTI)) %>% 
  dplyr::select(-TSH, -T3, -TT4, -T4U, -FTI)

#############################################
sick_t <- sick_t %>% mutate(sex = sex == "F",
                  on_thyroxine = on_thyroxine == "t",
                  query_on_thyroxine = query_on_thyroxine == "t",
                  on_antithyroid_medication = on_antithyroid_medication == "t",
                  sick = sick == "t",
                  pregnant = pregnant == "t",
                  thyroid_surgery = thyroid_surgery == "t",
                  I131_treatment  = I131_treatment  == "t",
                  query_hypothyroid  = query_hypothyroid  == "t",
                  query_hyperthyroid  = query_hyperthyroid  == "t",
                  lithium  = lithium  == "t",
                  goitre = goitre  == "t",
                  tumor = tumor  == "t",
                  psych = psych  == "t",
                  Class = Class == 'sick') %>%
  cbind(createDummyFeatures(sick_t$referral_source)) %>% 
  select(-referral_source) %>% as.matrix() %>% as.data.frame()

sick_test <- sick_t %>% filter(!ind) %>% select(-ind)
sick_t <- sick_t %>% filter(ind == 1) %>% select(-ind)

#############################################

```



```{r, echo=FALSE, message=FALSE, warning = FALSE, include=FALSE}

sick_t$Class <- as.factor(sick_t$Class)
  
task <- makeClassifTask(data = sick_t, target = "Class")
otask <- oversample(task, rate=5)

rdesc <- makeResampleDesc("CV", iters=5)
auprc.measure <- makeMeasure(id = "auprc",
            name = "AUPRC",
            properties = c('classif'),
            minimize = FALSE,
            best = 1,
            worst = 0,
            fun = function(task, model, pred, feats, extra.args) {
              auprc(pred$data$prob.1, pred$data$truth, 1)
            })



classifiers <- c(
  "classif.naiveBayes",
  "classif.binomial",
  "classif.rpart",
  "classif.ctree",
  "classif.C50",
  "classif.boosting",
  "classif.mlp",
  "classif.ada",
  "classif.gbm",
  "classif.glmboost",
  "classif.ranger",
  "classif.cforest")

models <- list()
omodels <- list()

for(classif in classifiers) {
  
  lr <- makeLearner(classif, predict.type = 'prob')
  
  m <- train(lr, task)
  p <- predict(m, newdata=sick_test)
  
  om <- train(lr, otask)
  op <- predict(om, newdata=sick_test)
  
  r <- suppressWarnings(resample(lr, task, rdesc, measures = list(auc, auprc.measure)))
  or <- suppressWarnings(resample(lr, otask, rdesc, measures = list(auc, auprc.measure)))
  
  models[[classif]] <- list(r$measures.test, p)
  omodels[[classif]] <- list(or$measures.test, op)
}

```

# Comparison between prediction accuracy of interpretable and non-interpretable models

I will compare five different interpretable models:

- naive biases

- logistic regression

- basic tree

- knn

- ctree

and back-box models:

- C50

- adabag boosting

- mlp 

- ada boost 

- gradient boosting machine 

- glm boost 

- random forest 

- cforest 


## Interpretable models

In this section we will examine prediction accuracy (via measuring AUC and AUPRC) of interpretable models. The following is not only a summary of the previous work, but an extension. Namely, due to imbalance of our classes, we performed oversampling, which generally has a capacity of increasing performance. 

### Naive Bayes
```{r, echo=FALSE, message=FALSE, warning = FALSE}
kable(models$classif.naiveBayes[[1]])
```

With oversampling:

```{r, echo=FALSE, message=FALSE, warning = FALSE}
kable(omodels$classif.naiveBayes[[1]])
```

Here we can see, that oversampling incresed AUPRC significantly. We must remember, that it doesn't mean that the model is that much better.

### Logistic Regression

```{r, echo=FALSE, message=FALSE, warning = FALSE}
kable(models$classif.binomial[[1]])
```

With oversampling:

```{r, echo=FALSE, message=FALSE, warning = FALSE}
kable(omodels$classif.binomial[[1]])
```

Again, oversampling gave better results.

### Tree

```{r, echo=FALSE, message=FALSE, warning = FALSE}
kable(models$classif.rpart[[1]])
```

With oversampling:

```{r, echo=FALSE, message=FALSE, warning = FALSE}
kable(omodels$classif.rpart[[1]])
```

Does the model changed? Let's see, this is a model without oversampling:

```{r, echo=FALSE, message=FALSE, warning = FALSE}
rpart.plot(train(makeLearner("classif.rpart", predict.type = 'prob'), task)$learner.model)
```

and with oversampling:

```{r, echo=FALSE, message=FALSE, warning = FALSE}
rpart.plot(train(makeLearner("classif.rpart", predict.type = 'prob'), otask)$learner.model)
```

Yes, it has changed. Oversampling then does change the model structure.

### The KNN

```{r, echo=FALSE, message=FALSE, warning = FALSE}
knn <- makeLearner("classif.kknn", predict.type = 'prob')
task_knn <- makeClassifTask(id="task_knn",data=sick_t %>% dplyr::select(log_TSH,
                                                   sqrt_T3,
                                                   sqrt_TT4,
                                                   sqrt_FTI,
                                                   Class,
                                                   query_hypothyroid), target="Class")

otask_knn <- oversample(task_knn, rate=5)

m <- train(knn, task_knn)
om <- train(knn, otask_knn)

p <- predict(m, newdata=sick_test %>% dplyr::select(log_TSH,
                                                   sqrt_T3,
                                                   sqrt_TT4,
                                                   sqrt_FTI,
                                                   Class,
                                                   query_hypothyroid))
op <- predict(om, newdata=sick_test %>% dplyr::select(log_TSH,
                                                   sqrt_T3,
                                                   sqrt_TT4,
                                                   sqrt_FTI,
                                                   Class,
                                                   query_hypothyroid))

r <- suppressMessages(resample(knn, task_knn, rdesc, measures = list(auc, auprc.measure)))
or <- suppressMessages(resample(knn, otask_knn, rdesc, measures = list(auc, auprc.measure)))

models[["classif.kknn"]] <- list(r$measures.test, p)
omodels[["classif.kknn"]] <- list(or$measures.test, op)

kable(models$classif.kknn[[1]])
```

With oversampling:

```{r, echo=FALSE, message=FALSE, warning = FALSE}
kable(omodels$classif.kknn[[1]])
```

The KNN works on a small subset of variables and it achieves very good results. 

### C-Tree

```{r, echo=FALSE, message=FALSE, warning = FALSE, fig.width=15, fig.height=15}
kable(models$classif.ctree[[1]])
```

With oversampling:

```{r, echo=FALSE, message=FALSE, warning = FALSE, fig.width=15, fig.height=15}
kable(omodels$classif.ctree[[1]])
```

It has very good results, but it's questionable if it's better than KNN. We would say, that it's not.

## Black-box Models

In this section I will examine the performance of back-box models.

### C50

```{r echo=FALSE, message=FALSE, warning=FALSE}
kable(models$classif.C50[[1]])
```

With oversampling:

```{r echo=FALSE, message=FALSE, warning=FALSE}
kable(omodels$classif.C50[[1]])
```

C50 shows comparable improvement with best interpretable classifiers.

### Adabag Boosting

```{r echo=FALSE, message=FALSE, warning=FALSE}
kable(models$classif.boosting[[1]])
```

With oversampling:

```{r echo=FALSE, message=FALSE, warning=FALSE}
kable(omodels$classif.boosting[[1]])
```

Adabag has the best results so far. It even maxed out auc in oversampling, which will not be achieved by other classifiers. 

### Ada Boost 

```{r echo=FALSE, message=FALSE, warning=FALSE}
kable(models$classif.ada[[1]])
```

With oversampling:

```{r echo=FALSE, message=FALSE, warning=FALSE}
kable(omodels$classif.ada[[1]])
```

Ada boost performed very well, scoring quite steady 0.95. It's not better than Adabag model, but it indicates that Ada/Adabag might be the way to go.

### Gradient Boosting Machine 

```{r echo=FALSE, message=FALSE, warning=FALSE}
kable(models$classif.gbm[[1]])
```

With oversampling:

```{r echo=FALSE, message=FALSE, warning=FALSE}
kable(omodels$classif.gbm[[1]])
```

Gradient Boosting Machine achieved very low scores, which is kind of surprise, it generally does well.

### GLM Boost 

```{r echo=FALSE, message=FALSE, warning=FALSE}
kable(models$classif.glmboost[[1]])
```

With oversampling:

```{r echo=FALSE, message=FALSE, warning=FALSE}
kable(omodels$classif.glmboost[[1]])
```

GLM boost didn't performed well.

### Random Forest 

```{r echo=FALSE, message=FALSE, warning=FALSE}
kable(models$classif.ranger[[1]])
```

With oversampling:

```{r echo=FALSE, message=FALSE, warning=FALSE}
kable(omodels$classif.ranger[[1]])
```

Random forest shows steady high auprc. It's an indicator of a good model.

### Cforest 

```{r echo=FALSE, message=FALSE, warning=FALSE}
kable(models$classif.cforest[[1]])
```

With oversampling:

```{r echo=FALSE, message=FALSE, warning=FALSE}
kable(omodels$classif.cforest[[1]])
```

It performed good, but not as good as other classifiers.

# Conclusions

Ada boost/Adabag boost performed best. Here's comparison of the performance of all classifiers on the test data set:

```{r, echo=FALSE, message=FALSE, warning=FALSE}
m <- as.data.frame(as.matrix(sort(sapply(models, function(tup) auprc(tup[[2]]$data[,3], tup[[2]]$data$truth, 1)))))
om <- as.data.frame(as.matrix(sort(sapply(omodels, function(tup) auprc(tup[[2]]$data[,3], tup[[2]]$data$truth, 1)))))
colnames(m) <- "auprc"
colnames(om) <- "auprc oversampled"
kable(cbind(m, om))

```

As we can see, Adabag boosting classifier achieved the best results, along with C50 and Ada. C-forest performed particularly bad, the opposite of what we predicted, while C50 turned out to work fine on a test sample. There can be seen a trend, that generally interpretable models perform worse than their back-box competitors, it's not a rule though.

The important thing about oversampling is that it did't change performance of the models. There is a bit of change in auprc in C50, however it just as well might be a statistical error.






