---
title: "Sick dataset analysis - blackbox"
author: "Michał Pastuszka"
date: "28 04 2020"
output: pdf_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
options(java.parameters = "-Xmx1g" )
library(OpenML)
library(dplyr)
library(tidyverse)
library(kableExtra)
library(visdat)
library(naniar)
library(DataExplorer)
library(caret)
library(PRROC)
library(corrplot)
library(glmnet)
library(mlr)
library(bartMachine)

```

# Loading dataset
```{r data, include = FALSE, cache=TRUE}

set.seed(10)

# download data
list_all_openml_dataset <- listOMLDataSets()

#sick dataset
openml_id <- 38 
data_name <- list_all_openml_dataset[list_all_openml_dataset[,'data.id'] == openml_id,'name']

dataset_openml <- getOMLDataSet(data.id = openml_id)
dataset_raw <- dataset_openml$data
target_column <- dataset_openml$target.features

```


# Preprocessing

```{r preprocessing, include = FALSE}

dataset <- dataset_raw %>% 
  # drop 'TBG' - it is an empty column:
  select(-TBG) %>%
  #drop 'TBG_measured' - it contains only one value:
  select(-TBG_measured)

```

```{r split}
train_indices <- read.csv('indeksy_treningowe.txt', sep = ' ', row.names = 1, header = TRUE)
dataset_train <- dataset[train_indices$x,]
dataset_test <- dataset[-train_indices$x,]
```

```{r drop_hypopituitary}
# variable hypopituary contains only one value in training set - we have to drop it
dataset_train <- dataset_train %>% 
  select(-hypopituitary)
dataset_test <- dataset_test %>% 
  select(-hypopituitary)
```

```{r imputation}
dataset_train$sex <- dataset_train$sex %>% replace_na('F')
dataset_test$sex <- dataset_test$sex %>% replace_na('F')
dataset_test$age <- dataset_test$age %>% replace_na(mean(dataset_train$age, na.rm = TRUE))
centerer_bag <- preProcess(dataset_train, method = c('bagImpute'))
dataset_train_bag <- predict(centerer_bag, dataset_train)
dataset_test_bag <- predict(centerer_bag, dataset_test)
```

# Blackbox model

We will fit a Ranger random forest using default parameters.

```{r ranger}
get_prauc <- function(responses, test_set){
  fg <- responses[test_set$Class == 'sick']
  bg <- responses[test_set$Class == 'negative']
  pr <- pr.curve(scores.class0 = fg, scores.class1 = bg, curve = T)
  plot(pr)
}
set.seed(10, "L'Ecuyer")
task <- makeClassifTask("ranger", data=dataset_train_bag, target = "Class")
learner_ranger <- makeLearner("classif.ranger", predict.type = 'prob')
model_ranger <- mlr::train(learner_ranger, task)
pred_ranger <- predict(model_ranger, newdata = dataset_test_bag)
probs <- pred_ranger$data$prob.sick

get_prauc(probs, dataset_test_bag)
```

It has completely outclassed the previous attempt using logistic regression. We Now we'll try to improve the performance even more tuning the hyperparamters. We will use 5 fold crossvalidation on the training set and mlrMBO
optimizer to tune mtry and min.node.size parameters.

```{r ranger_tune, cache=TRUE, warning=FALSE}
task <- makeClassifTask("ranger", data=dataset_train_bag, target = "Class")
learner_ranger <- makeLearner("classif.ranger", predict.type = 'prob')
ps <- makeParamSet(
  makeIntegerParam("mtry", lower = 1, upper = 26),
  makeIntegerParam("min.node.size", lower = 1, upper = 100)
)
ctrl <- makeTuneControlMBO()
rdesc <- makeResampleDesc("CV", iters = 5L, stratify = TRUE)
res <- tuneParams(learner_ranger, task, rdesc, auc, ps, ctrl, show.info = FALSE)
print(res)
learner_ranger_tune <- makeLearner("classif.ranger", predict.type = 'prob', par.vals = res$x)
model_ranger_tune <- mlr::train(learner_ranger_tune, task)
pred_ranger_tune <- predict(model_ranger_tune, newdata = dataset_test_bag)
probs <- pred_ranger_tune$data$prob.sick
get_prauc(probs, dataset_test_bag)
```

It scored even better than previously. It appears that using a blackbox model can drastically improve performance and reduce required work by sacrificing explainability and simplicity of the solution. It is worth noting, that a well tuned decision tree on this set can achieve similar results, while retaining explainability.