---
title: "Sick dataset analysis"
author: "Olaf Werner"
date: "17 04 2020"
output: pdf_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(OpenML)
library(dplyr)
library(tidyverse)
library(kableExtra)
library(visdat)
library(naniar)
library(DataExplorer)
library(dplyr)
library(readr)
library(mlr)
library(e1071)
library(PRROC)
```


```{r data, include = FALSE}

set.seed(10)

# download data
list_all_openml_dataset <- listOMLDataSets()

#sick dataset
openml_id <- 38 
data_name <- list_all_openml_dataset[list_all_openml_dataset[,'data.id'] == openml_id,'name']

dataset_openml <- getOMLDataSet(data.id = openml_id)
dataset_raw <- dataset_openml$data
target_column <- dataset_openml$target.features

```

```{r plot, warning=F}

DataExplorer::plot_histogram(dataset_raw)

DataExplorer::plot_bar(dataset_raw)

```

# Preprocessing

```{r preprocessing, include = FALSE}

dataset <- dataset_raw %>% 
  # drop 'TBG' - it is an empty column:
  select(-TBG)

```

```{r missings}

gg_miss_var(dataset, 
            show_pct = TRUE) + 
  ylim(0, 100) +
  labs(title = "Missing dataset",
       x = "Features",
       y = "Percent of missings")

# 80% train i 20% test data

# cross-validation

```

As we can see our dataset is very imbalanced on 3772 people only 231 are sick. So we are going to use AUC and AUPRC measures. Let us split our dataset to training and testing parts according to the file.

```{r warning=FALSE}
sick <- suppressMessages(read_csv("dataset_38_sick.csv"))
sick<-sick %>%
  mutate_if(is.character, list(~na_if(., "?")))
sick<-suppressMessages(type_convert(sick))
sick<-DataExplorer::drop_columns(sick,c("TBG","hypopituitary"))
sick<-sick %>% mutate_if(is.character,as.factor)
sick$Class<-sick$Class=="sick"
sick<-sick %>% mutate_if(is.logical,as.factor)

train_index<-suppressMessages(read_table2("indeksy_treningowe.txt"))
train_index<-train_index$y
train_sick<-sick[train_index,]
test_sick<-sick[-train_index,]
```

Because in mlr there is no AUPRC we create custom one using PRROC library.

```{r warning=FALSE }
require(PRROC)
AUPRC2<-function(task, model, pred, feats, extra.args){
  fg<-pred$data[pred$data$truth=="TRUE",]$prob.TRUE
  bg<-pred$data[pred$data$truth=="FALSE",]$prob.TRUE
  pr <- pr.curve(scores.class0 = fg, scores.class1 = bg, curve = T)
  pr$auc.integral
}
AUPRC<-makeMeasure(id = "auprc", minimize = FALSE,
                   properties = c("classif", "req.pred", "req.truth"),
                   best = 1, worst = 0, fun = AUPRC2)
```

In our dataset not only Class column is imbalanced, so we must consider dropping some columns out. Except hypopituitary because there is only 1 observation diffrent from the rest so we drop it right now.

# cforest without further cleaning

```{r cache=TRUE,warning=FALSE}
task<-makeClassifTask(data = train_sick, target = "Class")
rf_learner<-makeLearner("classif.cforest", predict.type = "prob")
rdesc = makeResampleDesc("CV", iters = 5)
r_rf = resample(rf_learner, task, rdesc, measures = list(mlr::auc, AUPRC),
                show.info = FALSE)

rf_model <- train(rf_learner, task)
prediction <- predict(rf_model, newdata = test_sick)

df = generateThreshVsPerfData(prediction, measures = list(fpr, tpr, mmce))
plotROCCurves(df)
res_rf=mlr::performance(prediction,list(mlr::auc,AUPRC))
```

```{r echo=FALSE}
score<-c(r_rf$aggr,res_rf)
names(score)<-c("auc.cross","auprc.cross","auc.test","auprc.test")
kable(data.frame(score),digits = 3)
score_rf<-score
```

Cforest is very complex and slow model but gives us very good results.
Our future models should try to be better than Cforest

# Rpart with 80% relevent columns

```{r cache=TRUE,warning=FALSE}
task<-makeClassifTask(data = train_sick, target = "Class")
rpart_learner<-makeLearner("classif.rpart", predict.type = "prob")
fv = generateFilterValuesData(task, method ="FSelectorRcpp_information.gain")
rdesc = makeResampleDesc("CV", iters = 5)
clean_task = filterFeatures(task, fval = fv, perc = 0.8)
r_rpart = resample(rpart_learner, clean_task, rdesc, measures = list(mlr::auc, AUPRC),
                   show.info = FALSE)
rpart_model <- train(rpart_learner, clean_task)
prediction <- predict(rpart_model, newdata = test_sick)
df = generateThreshVsPerfData(prediction, measures = list(fpr, tpr, mmce))
plotROCCurves(df)
res_rpart=mlr::performance(prediction,list(mlr::auc,AUPRC))

rpart.plot::rpart.plot(rpart_model$learner.model,roundint = FALSE)

```

```{r echo=FALSE}
score<-c(r_rpart$aggr,res_rpart)
names(score)<-c("auc.cross","auprc.cross","auc.test","auprc.test")
kable(data.frame(score),digits = 3)
score_rpart<-score
```

Results are not as good as Cforest but model iseft is very clear and simple.

# Rpart with generated dataset

```{r cache=TRUE,warning=FALSE}
train_sick_t<-train_sick[train_sick$Class=="TRUE",]
train_sick_f<-train_sick[train_sick$Class=="FALSE",]

random_t<-sapply(1:27,FUN = function(x){train_sick_t[sample(1:184,4000,replace = TRUE),x]})

random_f<-sapply(1:27,FUN = function(x){train_sick_f[sample(1:2832,1000,replace = TRUE),x]})

random_t<-data.frame(random_t)

random_f<-data.frame(random_f)

prediction <- predict(rf_model,newdata =  random_t)
response_t<-prediction$data$response
certainty<-abs(prediction$data$prob.TRUE-0.5)*2
sure_t<-certainty>0.3

prediction <- predict(rf_model,newdata =  random_f)
response_f<-prediction$data$response
certainty<-abs(prediction$data$prob.TRUE-0.5)*2
sure_f<-certainty>0.3

random_t$Class<-response_t

random_f$Class<-response_f

random<-rbind(random_t[sure_t,],random_f[sure_f,])

random<-rbind(random,train_sick)

task<-makeClassifTask(data = random, target = "Class")
rpart_learner<-makeLearner("classif.rpart", predict.type = "prob")
rdesc = makeResampleDesc("CV", iters = 5)
r_rpart_random = resample(rpart_learner, task, rdesc, measures = list(mlr::auc, AUPRC),show.info = FALSE)

rpart_model <- train(rpart_learner, task)
prediction <- predict(rpart_model, newdata = test_sick)

df = generateThreshVsPerfData(prediction, measures = list(fpr, tpr, mmce))
plotROCCurves(df)
res_rpart_random=mlr::performance(prediction,list(mlr::auc,AUPRC))

rpart.plot::rpart.plot(rpart_model$learner.model,roundint = FALSE)
```

```{r echo=FALSE}
score<-c(r_rpart_random$aggr,res_rpart_random)
names(score)<-c("auc.cross","auprc.cross","auc.test","auprc.test")
kable(data.frame(score),digits = 3)
score_rpart_random<-score
```

We generated additional data by splitting training dataset into parts with TRUE and FALSE classes then we generated random data in each part by resampling rows that means: for each column we were independently chosing random row and from such combination we got new observation. Classes of such obervations were determined by cforest we trained earlier. We also chose only such rows which cforest was certain of (probablity greater than 65%). Unfortunately we got worse model.

# Rpart tuned

```{r warning=FALSE}
learner<-makeLearner("classif.rpart", predict.type = "prob")
task<-makeClassifTask(data = train_sick, target = "Class")

dt_param <- makeParamSet(
  makeDiscreteParam("minsplit", values=seq(5,10,1)), 
  makeDiscreteParam("minbucket", values=seq(round(5/3,0), round(10/3,0), 1)),
  makeNumericParam("cp", lower = 0.01, upper = 0.05), 
  makeDiscreteParam("maxcompete", values=6), 
  makeDiscreteParam("usesurrogate", values=0),
  makeDiscreteParam("maxdepth", values=5) )

ctrl = makeTuneControlGrid()
rdesc = makeResampleDesc("CV", iters = 5L, stratify=TRUE)
(dt_tuneparam <- tuneParams(learner=learner,
                            resampling=rdesc,
                            measures=list(AUPRC,mlr::acc),
                            par.set=dt_param,
                            control=ctrl,
                            task=task,
                            show.info = FALSE) )

dtree <- setHyperPars(learner, par.vals = dt_tuneparam$x)

rdesc = makeResampleDesc("CV", iters = 5)
r_rpart_tuned = resample(dtree, task, rdesc, measures = list(mlr::auc, AUPRC),show.info = FALSE)

dtree_train <- train(learner=dtree, task=task)
prediction <- predict(dtree_train, newdata = test_sick)

df = generateThreshVsPerfData(prediction, measures = list(fpr, tpr, mmce))
plotROCCurves(df)
res_rpart_tuned=mlr::performance(prediction,list(mlr::auc,AUPRC))

rpart.plot::rpart.plot(dtree_train$learner.model,roundint = FALSE)
```

```{r echo=FALSE}
score<-c(r_rpart_tuned$aggr,res_rpart_tuned)
names(score)<-c("auc.cross","auprc.cross","auc.test","auprc.test")
kable(data.frame(score),digits = 3)
score_rpart_tuned<-score
```

Model after tuning performs a bit better than earlier, but is also more complicated.

# Summary

```{r echo=FALSE}
all_score<-rbind(data.frame(score=score_rf,type=c("auc.cross","auprc.cross","auc.test","auprc.test"),model="cforest"),
      data.frame(score=score_rpart,type=c("auc.cross","auprc.cross","auc.test","auprc.test"),model="rpart"),
      data.frame(score=score_rpart_tuned,type=c("auc.cross","auprc.cross","auc.test","auprc.test"),model="rpart_tuned"),
      data.frame(score=score_rpart_random,type=c("auc.cross","auprc.cross","auc.test","auprc.test"),model="rpart_random"))
```


```{r}
ggplot(all_score,aes(x=type,y=score,color=model))+geom_point()
```

Analizing this plot we can observe many interesting things. First of all our aproach with randomly generated data did not work. During cross validation on training data set it was getting similiar results to the cforest, but on test dataset it was getting worse results. This aproach may require more experiments and theoretical knownledge. Our tuned rpart was always bit worse during cross validation than default rpart, but on test dataset it was getting much better results in AUPRC and sligthy better AUC. 

