---
title: "Sick dataset analysis"
author: "Imię i nazwisko"
date: "25 03 2020"
output: pdf_document
---
  
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(OpenML)
library(dplyr)
library(tidyverse)
library(kableExtra)
library(visdat)
library(naniar)
library(DataExplorer)
library(Hmisc)
library(mlr3)
library(mlr3tuning)
library(auprc)
library(paradox)
```

# Intro

In this report I tried to keep everything simple. My data cleaning strategy was to delete all of the observations that might be problematic for my model as long as I am not deleting too many records of sick people. We are dealing with very unbalanced data and we can't afford to lose positive cases.


```{r data, include = FALSE}
set.seed(10)
# download data
list_all_openml_dataset <- listOMLDataSets()
#sick dataset
openml_id <- 38 
data_name <- list_all_openml_dataset[list_all_openml_dataset[,'data.id'] == openml_id,'name']
dataset_openml <- getOMLDataSet(data.id = openml_id)
dataset_raw <- dataset_openml$data
target_column <- dataset_openml$target.features
```

```{r plot, warning=F}
DataExplorer::plot_histogram(dataset_raw)
DataExplorer::plot_bar(dataset_raw)
DataExplorer::introduce(dataset_raw)
DataExplorer::plot_intro(dataset_raw)
```

# Preprocessing

```{r preprocessing, include = FALSE}
dataset <- dataset_raw %>% 
  # drop 'TBG' - it is an empty column:
  select(-TBG)
```

```{r missings, echo = FALSE}
gg_miss_var(dataset, 
            show_pct = TRUE) + 
  ylim(0, 100) +
  labs(title = "Missing dataset",
       x = "Features",
       y = "Percent of missings")

```

I want to delete all rows with missing data but also to control the number of positive cases in my data.

I also want to delete rows with some unrealistic values like 'age' above 400

```{r}
dataset <- dataset %>% mutate(age=replace(age, age>110 | age<0, NA))
```

In the future I am dropping all of the rows with NA so they won't be included in our model.

Next I want to see if there are any variables highly correlated that would cause our model to overfit.

```{r}
DataExplorer::plot_correlation(dataset)

DataExplorer::introduce(dataset)
DataExplorer::plot_intro(dataset)
# keeping track of positive cases before deleting rows
table(dataset$Class)
```

What I find out that there for sure are some columns with 0 variance. I also want to delete them. Other variables seem to be good.

First I will drop all the rows that currently have some missing values and see if we lost a lot of positive cases.

```{r}
dataset <- dataset %>% drop_na()

DataExplorer::introduce(dataset)
DataExplorer::plot_intro(dataset)
table(dataset$Class)
```

We didn't lose a lot of positive cases yet a lot of negatives ones which is good in case of inbalanced data.

Next I will delete all the columns with 0 variance. I will recognize them thanks to function `describe`.

```{r}
#describe(dataset)

dataset <- dataset %>% 
  select(-c("TBG_measured", "FTI_measured", "T4U_measured", "TT4_measured", "T3_measured", "TSH_measured" ))
```


# Modelling

First let's split data into train and test based on the file we got.

```{r}
train_idx <- read.table("indeksy_treningowe.txt", sep=" ", header = TRUE)$x
test_idx <- setdiff(1:3772, train_idx)
train <- dataset[train_idx,] %>% drop_na()
test <- dataset[test_idx,] %>% drop_na()

table(train$Class)
table(test$Class)
```

Classes ratios seem to be pretty similiar in both train and test set.

I want to use mlr3 for modelling and a simple decision tree that is easy to explain. I also used Dominik Rafacz's great package auprc that allows us to use this measure inside of mlr3. 

To use mlr3 I had to make sure it is using stratification when resampling our training set.

```{r}
measure_auc = msr('classif.auc')
measure = msr("classif.auprc")


task2 = TaskClassif$new(id = "classif_task",
                       backend = train,
                       target = "Class",
                       positive = 'sick')

resampling = rsmp("cv", folds = 5)

# r= resampling$instantiate(task)
# 
# prop.table(table(task$truth()))
# 
# prop.table(table(task$truth(r$train_set(1))))
# prop.table(table(task$truth(r$train_set(2))))
# prop.table(table(task$truth(r$train_set(3))))
# prop.table(table(task$truth(r$train_set(4))))
# prop.table(table(task$truth(r$train_set(5))))
```

We can see that each of the splits has similiar Class' ratios as the original train set. It means I can do the tuning using this resampling.


```{r message=FALSE, warning = FALSE, results = FALSE, cache = FALSE}
learner_rpart2 = lrn('classif.rpart', predict_sets = c("train", "test"), predict_type = 'prob')


tune_ps2 = ParamSet$new(list(
  ParamDbl$new("cp", lower = 0.001, upper = 0.1),
  ParamInt$new("minsplit", lower = 16, upper = 64),
  ParamInt$new("maxdepth", lower = 7, upper = 30)
))

evals20 = term("evals", n_evals = 3)

instance2 = TuningInstance$new(
  task = task2,
  learner = learner_rpart2,
  resampling = resampling,
  measures = measure,
  param_set = tune_ps2,
  terminator = evals20
)

tuner2 = tnr('random_search')
tuner2$tune(instance2)

learner_rpart2$param_set$values <- instance2$result$params
learner_rpart2$train(task2)

pred_train <- learner_rpart2$predict(task2)
pred_test <- learner_rpart2$predict_newdata(task2, newdata = test)

score_train <- pred_train$score(measure)
score_test <- pred_test$score(measure)

auc_train <- pred_train$score(measure_auc)
auc_test <- pred_test$score(measure_auc)
```

```{r}
results <- data.frame(
  'auc train' = auc_train,
  'auc_test' = auc_test,
  'auprc_train' = score_train,
  'auprc_test'= score_test
)

kable(results)
```

As we deal with unbalanced data `auprc` is much better measure to look at and it seems to be really good with our very simple approach.

We can also look at how easy to interpret this model is by looking at the tree visualization and variables' importances.

```{r cache = FALSE}
rpart.plot::rpart.plot(learner_rpart2$model,roundint=FALSE)

kable(learner_rpart2$importance())
```