---
title: "Sick dataset analysis"
author: "Szymon Maksymiuk"
date: "16 04 2020"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
    code_folding: hide
    number_sections: true
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(OpenML)
library(tidyverse)
library(kableExtra)
library(visdat)
library(naniar)
library(DataExplorer)

```


```{r data, include = FALSE}
set.seed(10)
dataset_openml <- getOMLDataSet(data.id = 38)
dataset_raw <- dataset_openml$data
test_index <- read.csv("index.txt", sep = " ")$x
```

# EDA

First let's look at basic information about those data. 

```{r}
summary(dataset_raw)
```
```{r}
DataExplorer::plot_missing(dataset_raw)
```

It is easy to spot straight away that one column, which is `TBG` contains 0 information. Therefore we will have it removed. Accordingly `TBG_measured` and `hypopituitary` have no information as well. 

```{r}
dataset_raw <- dataset_raw[,-which(names(dataset_raw)=="TBG")]
dataset_raw <- dataset_raw[,-which(names(dataset_raw)=="TBG_measured")]
dataset_raw <- dataset_raw[,-which(names(dataset_raw)=="hypopituitary")]
```

Before we move forward I would like to transform columns. I want factors with 2 levels to become a numeric vector with values 0 and 1.

```{r}
for (i in 1:ncol(dataset_raw)) {
  if (!is.null(levels(dataset_raw[[i]]))) {
    if (all(levels(dataset_raw[[i]]) %in% c("f", "t"))) {
      dataset_raw[i] <- as.numeric(ifelse(dataset_raw[i] == 't', 1, 0))
    }   
  }
}
dataset_raw$sex <- ifelse(dataset_raw$sex == "M", 1, 0)
dataset_raw$Class <- as.factor(ifelse(dataset_raw$Class == "sick", 1, 0))
```

```{r warning=FALSE}
DataExplorer::plot_bar(dataset_raw)
```

It is easy to spot a lot of difficulties that we will have to face. Data contains a lot of unbalanced categorical variables. The most extreme example is `lithium` that has 3541 entries of `f` value and only `18` of `t`. Another problem is that also target variable, which is `Class` is rather unbalanced. It can cause problems during the modeling stage.  



```{r warning=F}
DataExplorer::plot_histogram(dataset_raw)
```
```{r plot, warning=F}
DataExplorer::plot_qq(dataset_raw)
```


Distributions of continuous variables are quite pleasant. They are all a bit skewed but still quasi-normal what is good information. As exception stand `TSH` variable. Analysis of both histogram and qq plot shows a small anomaly.

```{r}
dataset_raw[dataset_raw$age>200,]
```

It is obviously a big mistake in data and therefore that row will be removed at the preprocessing stage.

# Preprocessing

Let's remove mistaken observation mentioned in the previous paragraph

```{r preprocessing, include = FALSE}
dataset_raw <- dataset_raw[-1365,]
```

Before we start proper preprocessing we have to split our data to test and train sets.

```{r}
dataset_test <- dataset_raw[-test_index,]
dataset_train <- dataset_raw[test_index,]
```

Now we need to handle missing values in observations. `mice` package will be used along with `pmm` method to impute missing values.

```{r}
library(mice)
m <- mice(dataset_train, method = "pmm", printFlag = FALSE)
dataset <- complete(m, 3)
```

# Modeling

Once we have our preprocessed data, it is hight time to start modeling part. `mlr` framework for modeling will be used. I planned on  using two types of white-box models:

* decision tree
* logistic regression.

And train hyperparameters using auprc as target measure. Unfortunately logistic regression model turned up to be terrible. Therefore for the rest of the work I will use a decision tree.

First I'm going to create a standard model that we want to improve.

```{r}
library("mlr")
library("auprc")
library("PRROC")

my_auprc_fun = function(task, model, pred, feats, extra.args) {
  prob <- pred$data$prob.1
  y_truth <- getPredictionTruth(pred)
  

  fg <- prob[y_truth == 1]
  bg <- prob[y_truth == 0]

  pr <- pr.curve(scores.class0 = fg, scores.class1 = bg, curve = T)
  auprc <- pr$auc.integral
  auprc
}

my_auprc = makeMeasure(
  id = "myauprc", name = "AUPRC",
  properties = c("classif", "classif.multi", "req.pred", "req.truth"),
  minimize = FALSE, best = 1, worst = 0,
  fun = my_auprc_fun
)

task <- makeClassifTask(data = dataset, target = "Class")
lrn_tree <- makeLearner("classif.rpart", predict.type = "prob")

model_tree <- train(lrn_tree, task)

p_tree <- predict(model_tree, newdata = dataset_test)

performance(p_tree, list(auc, my_auprc))
```

```{r eval=FALSE}
library(mlrMBO)
par.set <- makeParamSet(
  makeIntegerParam("minsplit", 3, 40),
  makeIntegerParam("minbucket", 2, 40),
  makeNumericParam("cp", -10, -3, trafo = function(x) 2^x),
  makeIntegerParam("maxdepth", 1, 30)
)
cv <- makeResampleDesc("CV", iters = 5L)
ctrl <- makeMBOControl()
ctrl <- setMBOControlTermination(ctrl, iters = 100)
tune_ctrl <- makeTuneControlMBO(mbo.control = ctrl)
res_tree <- tuneParams(lrn_tree, task, cv, par.set = par.set, control = tune_ctrl)
```

Tuning has slightly imporved our model. Here are results.

```{r}
lrn_tree <- makeLearner("classif.rpart", predict.type = "prob", par.vals = list(minsplit = 25,
                                                                                minbucket = 7,
                                                                                cp = 0.00108))

model_tree <- train(lrn_tree, task)

p_tree <- predict(model_tree, newdata = dataset_test)

performance(p_tree, list(auc, my_auprc))
```
